---
title: "Final Assignment"
author: "Justin Gibson"
date: "31/03/2022"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
list=ls(all=TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library("writexl")
library(ggplot2)
library(gbm)
library(glmnet)
library(randomForest)
library(rpart)
library(rpart.plot)
library(tidyverse)
library(tibble)

# cd ~/Desktop/Final\ Project
# git commit -am "Message"
# git push
```

```{r setup 2, include=FALSE}
# Importing CSV after formatting modifications in Excel
NHL.df = data.frame(read_csv("/Users/justingibson/Desktop/Adv Topics Fin/NHL Final Project/NHL_df.csv"))
attach(NHL.df)
```

```{r introduction}
# If there are any missing salaries for any NHL player, this gets rid of them
NHL.df = NHL.df[complete.cases(Salary),]

# removing every player on a entry level contract
NHL.df = NHL.df %>%
  filter(Salary >= 950000)

# Plotting every NHL player and their respected salary
player.num.vec = c(1:463)
sal.intro.plot = ggplot(NHL.df, aes(x=player.num.vec, y=Salary)) + geom_point()
sal.intro.plot
```

This plot shows the importance of this research question. It is evident that NHL hockey players' salaries vary greatly amongst NHL players and there is a lot more research we can do to understand why a hockey player is being paid a given salary.

```{r introduction 1}
# Log transformation of Salaries
NHL.df$Salary = log(NHL.df$Salary)

# New plot of NHL players' salaries
sal.intro.plot.log = ggplot(NHL.df, aes(x=player.num.vec, y=NHL.df$Salary)) + geom_point()
sal.intro.plot.log
```

```{r linear regression}
# Linear regression on all the data to see if any variables can explain all players
set.seed(1)
lin.reg = lm(NHL.df$Salary ~ Age + Contract.Length + t_on_ice + assists + goals + pim + shots + games + hits + pp_goals + pp_points + pp_t_on_ice + even_t_on_ice + pen_min + face_off_perc + shot_perc + game_win_goals + over_t_goals + short_hand_goals + short_hand_points + short_hand_t_on_ice + blocked + plus_minus + points + shifts, data = NHL.df)

summary(lin.reg)

```

Omitted logistic variables, but kept all numeric variables. This provided very interesting results, we now have a better understanding of what variables have explanatory power for the independent variable, salary. The variables: Age, contract.length, assists, games, and face_off_perc all have explanatory power. Further linear regression will be completed below to determine what variables have the most explanatory power by removing variables that have little explanatory power.

One variable was removed if two variables have strong a correlation, i.e (shots & shot_perc, t_on_ice & shifts). Other variables were removed as well that had high p-values.

```{r lin reg explanatory power}
#Removed variables, hoping to find variables with more explanatory power
lin.reg1 = lm(NHL.df$Salary ~ Age + Contract.Length + assists + goals + pp_points + shots + games + face_off_perc + short_hand_points + plus_minus + shifts + pp_t_on_ice, data = NHL.df)

summary(lin.reg1)
```

Shots, face_off_perc, and plus_minus are removed as they all are not significant

```{r}
# More variables were taken out, notice the increased r^2 below
lin.reg2 = lm(NHL.df$Salary ~ Age + Contract.Length + assists + goals + games + short_hand_points + shifts + pp_t_on_ice, data = NHL.df)

summary(lin.reg2)
```

Not surprisingly contract length is correlated with Salaries. It is understandable that NHL teams want to sign players for long contracts if they are going to invest in them and their performance.

```{r}
lin.reg3 = lm(NHL.df$Salary ~ Age + Contract.Length + assists + games + pp_t_on_ice, data = NHL.df)

summary(lin.reg3)
```

```{r filtering by position}
# Using dplyr to filter for all forward position NHL players (grouping all forwards)
forwards.vec = c("LW", "RW", "C")

forwards.df = NHL.df %>%
  filter(Position %in% forwards.vec)

# Using dplyr to filter for all defense position NHL players (grouping all defense)
defense.vec = c("D")

defense.df = NHL.df %>%
  filter(Position %in% defense.vec)

```

```{r training and test set for forward and defense players}
set.seed(2)
# Train and test set for all forwards players
samp.set.for = sample(nrow(forwards.df), 0.75*nrow(forwards.df))
train.for = forwards.df[samp.set.for,]
test.for = forwards.df[-samp.set.for,]


# Train and test set for all defense players
samp.set.def = sample(nrow(defense.df), 0.75*nrow(defense.df))
train.def = defense.df[samp.set.def,]
test.def = defense.df[-samp.set.def,]


```


```{r lin reg forwards}
# Linear regression process on all forwards
lin.reg.for = lm(train.for$Salary ~ Age + Contract.Length + t_on_ice + assists + goals + pim + shots + games + hits + pp_goals + pp_points + pp_t_on_ice + even_t_on_ice + pen_min + face_off_perc + shot_perc + game_win_goals + over_t_goals + short_hand_goals + short_hand_points + short_hand_t_on_ice + blocked + plus_minus + points + shifts, data = train.for)

summary(lin.reg.for)
```
Keeping all performance measures that seem applicable to forward players. Removing very p-value variables

```{r lin reg forwards 1}
# Got rid of Contract length + Plus_minus + shot_perc + shifts + pen_min + points
lin.reg.for.1 = lm(train.for$Salary ~ Age + t_on_ice + assists + goals + pim + shots + games + hits + pp_goals + pp_points + pp_t_on_ice + even_t_on_ice + face_off_perc + game_win_goals + over_t_goals + short_hand_goals + short_hand_points + short_hand_t_on_ice + blocked, data = train.for)

summary(lin.reg.for.1)
```


```{r}
# Got rid of games, goals, pim, game win goals, short hand goals
lin.reg.for.2 = lm(train.for$Salary ~ Age + t_on_ice + assists + shots + games + hits + pp_goals + pp_points + pp_t_on_ice + even_t_on_ice + pen_min + face_off_perc + over_t_goals + short_hand_t_on_ice + blocked + points, data = train.for)

summary(lin.reg.for.2)

```

```{r}
# Got rid of shots, pp_goals, pen_min
lin.reg.for.3 = lm(train.for$Salary ~ Age + t_on_ice + assists + games + hits + pp_points + pp_t_on_ice + even_t_on_ice + face_off_perc + over_t_goals + short_hand_t_on_ice + blocked + points, data = train.for)

summary(lin.reg.for.3)
```

```{r}
# Got rid of blocked, points, hits
lin.reg.for.4 = lm(train.for$Salary ~ Age + t_on_ice + assists + games + pp_points + pp_t_on_ice + even_t_on_ice + face_off_perc + over_t_goals + short_hand_t_on_ice, data = train.for)

summary(lin.reg.for.4)
```

```{r}
# Predicting salary on test set
for.pred.test = predict(lin.reg.for.4, newdata = test.for)

# MSE on Linear Regression for forward players
MSE.for = mean((for.pred.test - test.for$Salary)^2)
MSE.for

# Predicting salary for each player in the forwards data frame
Linear.Prediction = predict(lin.reg.for.4, newdata=forwards.df)

chg.salary.for.df = data.frame(forwards.df, Linear.Prediction)

# Creating a change in salary column in new data frame
chg.salary.for.df = chg.salary.for.df %>%
  mutate(Change.in.Salary.Linear = forwards.df$Salary - Linear.Prediction)
```

```{r}
# Linear regression process for defense players
lin.reg.def = lm(train.def$Salary ~ Age + Contract.Length + t_on_ice + assists + goals + pim + shots + games + hits + pp_goals + pp_points + pp_t_on_ice + even_t_on_ice + pen_min + face_off_perc + shot_perc + game_win_goals + over_t_goals + short_hand_goals + short_hand_points + short_hand_t_on_ice + blocked + plus_minus + points + shifts, data = train.def)

summary(lin.reg.def)
```

```{r}
# Removed contract length, pen_min, face_off_perc, points, goals
lin.reg.def.1 = lm(train.def$Salary ~ Age + t_on_ice + assists + pim + shots + games + hits + pp_goals + pp_points + pp_t_on_ice + even_t_on_ice + shot_perc + game_win_goals + over_t_goals + short_hand_goals + short_hand_points + short_hand_t_on_ice + blocked + plus_minus + shifts, data = train.def)

summary(lin.reg.def.1)
```

```{r}
# Removed over_t_goals, blocked, pp_goals, game_win_goals, short_hand_goals, short_hand_points
lin.reg.def.2 = lm(train.def$Salary ~ Age + t_on_ice + assists + pim + shots + games + hits + pp_points + pp_t_on_ice + even_t_on_ice + shot_perc + short_hand_t_on_ice + plus_minus + shifts, data = train.def)

summary(lin.reg.def.2)
```

```{r}
# Removed assists, pim, hits, shot_perc
lin.reg.def.3 = lm(train.def$Salary ~ Age + t_on_ice + shots + games + pp_points + pp_t_on_ice + even_t_on_ice + short_hand_t_on_ice + plus_minus + shifts, data = train.def)

summary(lin.reg.def.3)
```

```{r}
# Predicting salary for each defense player in test set
def.pred.test = predict(lin.reg.def.3, newdata = test.def)

# Mean squared error for linear regression of defense players
MSE.def = mean((def.pred.test - test.def$Salary)^2)
MSE.def

# Predicting salary for each player in the defense player
Linear.Prediction = predict(lin.reg.def.3, newdata=defense.df)

chg.salary.def.df = data.frame(defense.df, Linear.Prediction)

# Creating a change in salary column in new data frame
chg.salary.def.df = chg.salary.def.df %>%
  mutate(Change.in.Salary.Linear = defense.df$Salary - Linear.Prediction)

```

```{r}
# Creating a data frame of just player name, position and salary for each position

forwards.df.conc = chg.salary.for.df %>%
    select(Player,Position,Salary,Linear.Prediction,Change.in.Salary.Linear)  

defense.df.conc = chg.salary.def.df %>%
    select(Player,Position,Salary,Linear.Prediction,Change.in.Salary.Linear) 

set.seed(3)

  # Boost reg forwards
boost.NHL.for = gbm(train.for$Salary ~ Age + Contract.Length + t_on_ice + assists + goals + pim + shots + games + hits + pp_goals + pp_points + pp_t_on_ice + even_t_on_ice + pen_min + shot_perc + game_win_goals + over_t_goals + short_hand_goals + short_hand_points + short_hand_t_on_ice + blocked + plus_minus + points + shifts, data=train.for, distribution = "gaussian", n.trees = 1000, interaction.depth = 4)
  
  summary(boost.NHL.for)
  
  # Predicting the boost regression of forward players on test set
  boost.NHL.pred.for = predict(boost.NHL.for, newdata=test.for, n.trees = 1000, interaction.depth = 4)
  
  # Mean squared error for boost regression of forward players
  boost.MSE = mean((boost.NHL.pred.for - test.for$Salary)^2)
  boost.MSE
  
  # Predicting salary for each player in the forward player
  Boost.Prediction = predict(boost.NHL.for, newdata=forwards.df, n.trees = 1000, interaction.depth = 4)

  Forwards.Salaries.df = data.frame(forwards.df.conc, Boost.Prediction)

  # Creating a change in salary column in new data frame
  Forwards.Salaries.df = Forwards.Salaries.df %>%
   mutate(Change.in.Salary.for.Boost = forwards.df$Salary - Boost.Prediction)


  # Boost reg defense
  boost.NHL.def = gbm(train.def$Salary ~ Age + Contract.Length + t_on_ice + assists + goals + pim + shots + games + hits + pp_goals + pp_points + pp_t_on_ice + even_t_on_ice + pen_min + shot_perc + game_win_goals + over_t_goals + short_hand_goals + short_hand_points + short_hand_t_on_ice + blocked + plus_minus + points + shifts, data=train.def, distribution = "gaussian", n.trees = 1000, interaction.depth = 4)
  
  summary(boost.NHL.def)
  
  # Predicting boost regression on test set of defense players
  boost.NHL.pred.def = predict(boost.NHL.def, newdata=test.def, n.trees = 1000, interaction.depth = 4)
  
  # Mean squared error of boost regression on test set
  boost.MSE.def = mean((boost.NHL.pred.def - test.def$Salary)^2)
  boost.MSE.def
  
  # Predicting salary for each player in the defense player
  Boost.Prediction = predict(boost.NHL.def, newdata=defense.df, n.trees = 1000, interaction.depth = 4)

  Defense.Salaries.df = data.frame(defense.df.conc, Boost.Prediction)

  # Creating a change in salary column in new data frame
  Defense.Salaries.df = Defense.Salaries.df %>%
   mutate(Change.in.Salary.for.Boost = defense.df$Salary - Boost.Prediction)

```

By preforming a boost regression on the entire NHL data frame, it is evident that Contract Length has the most explanatory power with respect to Salary. The runner up is pp_t_on_ice which is power play time on ice. This is a surprising result. This tells a different story then the simple linear regression preformed above where power play time on ice was less significant then many other variables.


```{r merging everything back into one table}
# Preparing final data
Forwards.Salaries.df = Forwards.Salaries.df %>%
  mutate(Net.Change = mean(Change.in.Salary.Linear + Change.in.Salary.for.Boost)) %>%
  mutate(Final.Salary = Forwards.Salaries.df$Salary + Net.Change)

Defense.Salaries.df = Defense.Salaries.df %>%
  mutate(Net.Change = mean(Change.in.Salary.Linear + Change.in.Salary.for.Boost)) %>%
  mutate(Final.Salary = Defense.Salaries.df$Salary + Net.Change)

Final.data = rbind(Forwards.Salaries.df, Defense.Salaries.df)
Final.data = Final.data %>%
  arrange(desc(Final.Salary))

View(Final.data)

```






